---
title: "Reproducible Research: Peer Assessment 1"
author: José Oliver Segura
output: 
html_document:
keep_md: true
---


## Loading and preprocessing the data
```{r output=F}
options(digits=12)
```


Assuming **cwd** is set to the directory where data is located, we can load data directly from the zip file:

```{r}
content <- unz("activity.zip","activity.csv")
dataset <- read.csv(content,stringsAsFactors = T)
```

We can check data is loaded and looks fine:
```{r}
head(dataset)
```
```{r}
summary(dataset)
```

## What is mean total number of steps taken per day?

Since we have to respond some questions about total steps per day, we can create a dataset of aggregated data:

```{r}
agg.date <- aggregate(steps ~ date, dataset, sum)
head(agg.date)
```

Next, we can plot the histogram of total steps per day:
```{r}
hist(agg.date$steps,breaks=10,xlab="steps",main="Total steps per day histogram",col="gray")
```

As demanded, we can find out the mean and the median for the total number of steps for each day:

```{r}
summary(agg.date$steps)
```

We can observe that the mean is `r sprintf("%.4f",mean(agg.date$steps))` and the median is `r sprintf("%d",median(agg.date$steps))`.

## What is the average daily activity pattern?

Let's make a time series plot of the 5 minute interval and the average number of steps taken across all days. Since we need also to solve another questions, we can aggregate now by interval:

```{r}
agg.interval <- aggregate(steps ~ interval,dataset,mean)
head(agg.interval)
```

Now we can plot the data:
```{r}
plot(agg.interval,type="l",main="Average steps per interval")
```

As can be seen in the plot, the 5-minute interval with the maximum number of steps (on average) is somewhere near the 1000th interval, we can find it exactly:

```{r}
index <- which.max(agg.interval$steps)
agg.interval$interval[index]
```

## Imputing missing values

Let's find out how many missing values do we have:

```{r}
sapply(names(dataset),function(n) {sum(is.na(dataset[n]))})
```

Looks that the only column with missing values is "steps".



## Are there differences in activity patterns between weekdays and weekends?
